---
title: "Principal Component Analysis"
author: "Derek Nguyen"
date: "2020-02-04"
output: html_document
tags:
  - statistical
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(tidyr)
library(janitor)
library(ggfortify)
library(here)
```


### Introduction 

Here we explore the a dataset provided by, *[Google Earth Engine](https://developers.google.com/earth-engine/datasets/catalog)*, that provides various environmental and climatic data on countries throughout the world. For this particular project, variables observed are: annual mean temperature, annual mean rainfall, accessibility to cities, tree canopy cover, cloudiness, elevation, and cropland cover.  

```{r, warning = FALSE, message = FALSE}
world_env <- read_csv(here::here("csv", "world_env_vars.csv"))
```

```{r, warning = FALSE, message = FALSE}
world_env_tidy <- world_env %>% 
  clean_names() %>% 
  select(cropland_cover,
         tree_canopy_cover, 
         accessibility_to_cities, 
         elevation, 
         rain_mean_annual,
         temp_mean_annual,
         cloudiness) %>% 
  drop_na()
  
```

```{r}

world_env_pca <- prcomp(world_env_tidy, scale = TRUE)
```

```{r, message = FALSE, echo = FALSE }


world_env_biplot <- autoplot(world_env_pca,
                             colour = NA,
                             loadings.label = TRUE,
                             loadings.label.size = 3,
                             loadings.label.colour = "black",
                             loadings.label.repel = TRUE) +
  theme_minimal() + 
  scale_y_continuous(limits = c(-.2, .2)) +
  labs(title = "PCA of Country-Level Environmental and Climatic Variables")

world_env_biplot
```



### Summary
- Mean annual rainfall and tree canopy cover is most positively correlated. This correlation makes sense because rainfall provides nutrients for trees and tree canopy cover would grow. 
- Mean annual rainfall is also positively correlated with cloudiness. This seems valid because cloudy days mean higher potential for rainfall. 
- Mean annual temperature and cropland cover are inversely correlated. 
- Mean annual temperature and elevation are inversely correlated. This makes sense because as elevation increases, temperature goes down and vice versa.


























